---
output: github_document
---

<!-- README.md is generated from the README.Rmd. Please edit that file -->

# Ontario epidemic curve reconstruction
This is the code to replicate the results in Section 3 of the manuscript. 

Start by running the script **1_1_wastewatermodel_ON.R** to fit the wastewater model. For more information on the wastewater model, please see https://doi.org/10.1093/jrsssc/qlae073 and https://github.com/emilysomerset/wastewater_paper_code. 

For the remainder of the analysis, you will need the following libraries. The R version used for this analysis and the package versions are listed below.

```{r setup, warning=FALSE, message=FALSE}
# > sessionInfo()
# R version 4.4.1 (2024-06-14)
# Platform: aarch64-apple-darwin20
# Running under: macOS Monterey 12.4

# R version 4.4.2 (2024-10-31)
# Platform: x86_64-pc-linux-gnu
# Running under: Ubuntu 22.04.5 LTS

library(dplyr) # dplyr_1.1.4
library(TMB) # TMB_1.9.16
library(magrittr) # magrittr_2.0.3
library(reshape2) # reshape2_1.4.4
library(gridExtra) # gridExtra_2.3
library(cowplot) # cowplot_1.1.3
library(lubridate) # lubridate_1.9.4
library(ggplot2) # ggplot2_3.5.1
```

You will need to source the following functions, provided in this folder. 
```{r, warning=FALSE, message=FALSE}
source('../functions_general/prep_data_covid_with_fittedwastewater.R')
source('../functions_general/process_results_epidemic.R')
```

Compile and load the following C++ file
```{r, warning=FALSE, message=FALSE}
compile(file="../section_3/cpp/model4.cpp")
try(dyn.unload(dynlib("../section_3/cpp/model4")),silent = TRUE)
dyn.load(dynlib("../section_3/cpp/model4"))
```

## Load case data and prep data

```{r, warning=FALSE, message=FALSE}
load("../section_3/data/work_d_toronto_2024_08_08.RData") # Public health ontario covid cases
load("../section_3/data/work_d_pho_testing_2024_10_07.RData") # Public health ontario testing
load(file="../section_3/results_model_ON_phachist.RData") # from the wastewater model

work_d_toronto <- work_d_toronto %>% 
  left_join(work_d_testing %>% 
              dplyr::select(week_end_date, total_number_of_tests) %>% 
              mutate(week_end_date = ymd(week_end_date)), 
            by = c("earliest_week_end_date"="week_end_date"))

# bring in hospital data too

load("../section_3/data/work_d_pho_hosp_cases_2024_10_07.RData")
load("../section_3/data/work_d_toronto_hosp_cases_2024_09_05.RData")

work_d_toronto <- work_d_toronto %>% 
  left_join(work_d_hosp2 %>% 
              filter(outcome == "COVID-19 hospital admissions (up to January 20, 2024)") %>% 
              dplyr::select(week_end_date, number) %>% 
              mutate(week_end_date = ymd(week_end_date)) %>% 
              rename("admissions" = number), 
            by = c("earliest_week_end_date"="week_end_date")) %>% 
    left_join(work_d_hosp2 %>% 
                filter(outcome == "COVID-19 deaths") %>% 
                dplyr::select(week_end_date, number,population) %>% 
                mutate(week_end_date = ymd(week_end_date)) %>% 
                rename("deaths" = number),
            by = c("earliest_week_end_date"="week_end_date")) %>% 
      left_join(work_d_hosp %>% 
              dplyr::select(episode_week, hospitalized_cases) %>% 
              mutate(episode_week = ymd(episode_week)), 
            by = c("week_start_date"="episode_week"))

# These weights are from 
# https://www.toronto.ca/wp-content/uploads/2023/10/8e9f-PublicHealthWastewaterSurveillanceTechNotes.pdf
weights_datadic = data.frame(site_id = c("TAB","THC","THU","TNT"), 
                             weight = c(0.499/(0.177+0.232+0.064+0.499),
                                        0.177/(0.177+0.232+0.064+0.499),
                                        0.232/(0.177+0.232+0.064+0.499),
                                        0.064/(0.177+0.232+0.064+0.499)))

data_foranalysis_full <- prep_data(case_data = work_d_toronto,
                                   y_var = "number_of_cases",
                                   results = results,
                                   AR=TRUE,
                                   weight_ratio = TRUE,
                                   weights_datadic = weights_datadic)

pop <- work_d_toronto %>% 
  filter(earliest_week_end_date < mdy("09-30-2020")) %>% slice(1) %$% population[1]

```




```{r, cache=TRUE, echo=FALSE}
tmp = Reduce(rbind, lapply(as.list(1:3000),function(i){data_foranalysis_full$analysis_d[[i]]$version = i; data_foranalysis_full$analysis_d[[i]]}))
tmp <- tmp %>% 
  group_by(earliest_week_end_date,y) %>% 
  summarise(ratio_v_u_fixed_lwr = quantile(ratio_v_u_fixed,0.025),
            ratio_v_u_fixed_upr = quantile(ratio_v_u_fixed,0.975),
            ratio_v_u_fixed_med = quantile(ratio_v_u_fixed,0.5),
            ratio_lwr = quantile(ratio,0.025),
            ratio_upr = quantile(ratio,0.975),
            ratio_med = quantile(ratio,0.5))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
tmp %>%ungroup() %>%  mutate(lagged_case = lag(y)) %>% 
  ggplot(aes(earliest_week_end_date, y/lagged_case))+
  geom_line(aes(earliest_week_end_date,ratio_v_u_fixed_med), alpha = 0.5)+  
  geom_ribbon(aes(earliest_week_end_date,ymax = ratio_v_u_fixed_upr, ymin = ratio_v_u_fixed_lwr), alpha = 0.5)+  
  geom_point(col="red")+
  geom_line(col ="red")+
  theme_bw()+ 
  scale_y_continuous(name = "Effective reproduction number",
                     breaks = scales::pretty_breaks(n=10))+
  scale_x_date(name = "Epidemiological week",
                     breaks = scales::pretty_breaks(n=10))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
tmp %>%ungroup() %>%  mutate(lagged_case = lag(y)) %>% 
  ggplot(aes(earliest_week_end_date, y/lagged_case))+
  geom_line(aes(earliest_week_end_date,ratio_med), alpha = 0.5)+  
  geom_ribbon(aes(earliest_week_end_date,ymax = ratio_upr, ymin = ratio_lwr), alpha = 0.5)+  
  geom_point(col="red")+
  geom_line(col ="red")+
  theme_bw()+ 
  scale_y_continuous(name = "Effective reproduction number",
                     breaks = scales::pretty_breaks(n=10))+
    scale_x_date(name = "Epidemiological week",
                     breaks = scales::pretty_breaks(n=10))
```

## Fit model

```{r}
# This model is fit using the code in 2_1_popweighted_fixed_pred.R
# We load the model that was fit using the full dataset (j = 126)

load("../section_3/results_pred_popweighted_fixed/results_126.RData")
```

## Process the results

```{r, message=FALSE, warning=FALSE}
results <- process_results(data_foranalysis = data_foranalysis_full,
                MI_models = MI_models,
                mean_adj = 2.57/100*pop, 
                sd_adj = 0.47/100*pop)
```

## Plot the results

### Ascertainment probability

```{r, echo=FALSE, warning=FALSE, message=FALSE, out.height="50%"}

results

ggplot(results ,  aes(earliest_week_end_date, p_med))+
  geom_line()+
  # geom_point()+
  geom_ribbon(aes(ymin = p_lwr, ymax = p_upr), alpha = 0.3)+
  theme_bw()+
  xlab("Epidemiological week")+ 
  scale_y_continuous(name = "Ascertainment probability", breaks = scales::pretty_breaks(n=10))+
  geom_line(aes(earliest_week_end_date,total_number_of_tests/max(total_number_of_tests)))

```

### Weekly new infection counts

```{r, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(results, aes(earliest_week_end_date, z_med))+
  geom_line()+
  geom_line(aes(earliest_week_end_date, y), col = "red")+
  geom_ribbon(aes(ymin = z_lwr, ymax = z_upr), alpha = 0.3)+
  theme_bw()+ 
  xlab("Epidemiological week")+ 
  scale_y_continuous(name = "New infections", breaks = scales::pretty_breaks(n=10))
```


### Cumulative infections

```{r, echo=FALSE, warning=FALSE, message=FALSE}

results %>% 
  filter(earliest_week_end_date<= ymd("2021-04-03") & earliest_week_end_date >= ymd("2021-03-27")) %>% 
  dplyr::select(earliest_week_end_date, 
                z_cumsum_noadj_delta_med,
                z_cumsum_noadj_delta_lwr,
                z_cumsum_noadj_delta_upr,
                number_of_cases_cumsum_delta) %>% 
  mutate(z_cumsum_noadj_delta_med=z_cumsum_noadj_delta_med/pop*100,
         z_cumsum_noadj_delta_lwr=z_cumsum_noadj_delta_lwr/pop*100,
         z_cumsum_noadj_delta_upr=z_cumsum_noadj_delta_upr/pop*100,
         number_of_cases_cumsum_delta = number_of_cases_cumsum_delta/pop*100)

ggplot(results %>% filter(earliest_week_end_date<"2021-12-01"), aes(earliest_week_end_date, z_cumsum_noadj_delta_med/pop*100))+
  geom_line()+
  geom_ribbon(aes(ymin = z_cumsum_noadj_delta_lwr/pop*100, ymax = z_cumsum_noadj_delta_upr/pop*100), alpha = 0.3)+
  theme_bw()+ 
  xlab("Epidemiological week")+ 
  scale_y_continuous(name = "Cumulative Incidence (%)", breaks = scales::pretty_breaks(n=10))+
  geom_errorbar(data= data.frame(earliest_week_end_date=ymd("2021-03-31"),
                                 lwr = 3.74,
                                 upr = 7.27), 
                aes(x=earliest_week_end_date,ymax =upr, ymin = lwr),inherit.aes = FALSE,size=0.5,width = 10)+
     geom_point(data = data.frame(earliest_week_end_date=ymd("2021-03-31"),y= 5.38 ),
             aes(earliest_week_end_date, y),size= 0.5)+
  geom_line(aes(earliest_week_start_date, number_of_cases_cumsum_delta/pop*100), col = "red")
```
